---
title: "Mesenchymal analysis by eye"
author: "Jason T. Serviss"
date: "30/06/2017"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    theme: readable
editor_options: 
  chunk_output_type: console
---

```{r loadLibraries, message=FALSE}
packages <- c(
    "mesenchymalSubsetAnalysis", 
    "printr",
    "ggthemes",
    "tidyverse",
    "mclust",
    "viridis"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)
```

```{r setup}
tsne <- my.tsne.2d.full.top2000.log
counts <- counts.nodups
counts.norm <- countsNorm(counts)
counts.log <- log2(counts.norm)

groups.mesenchymal <- groups.mesenchymal.n
groups.fetal <- groupNames(groups.fetal)

mes.tsne <- tsne[groups.fetal == "Mesenchymal", ]
```

### Plot all cell types 

```{r plotAllCellTypes, fig.align='center', fig.height=8, fig.width=10}
data.frame(tsne, class = groups.fetal) %>%
  as_tibble() %>%
  ggplot() +
  geom_point(aes(X1, X2, colour = class)) +
  theme_few() +
  scale_colour_ptol()
```

### Plot the mesenchymal cell subtypes
Subtypes classified via Mclust.

```{r, fig.align='center', fig.height=8, fig.width=10}
mes.tsne %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  inner_join(sampleClasses, by = "sample") %>%
  as_tibble() %>%
  filter(class != "UndefinedMesenchymal") %>%
  ggplot() +
  geom_point(aes(V1, V2, colour = class)) +
  theme_few() +
  scale_colour_ptol() +
  labs(x = "t-SNE dim 1", y = "t-SNE dim 2") +
  guides(colour = guide_legend(title = "Cell subtype"))
```

### Select genes that identify each mesenchymal subset.
Genes specific expression for each cell type (including mesenchymal subclasses)
was then analyzed using the Kolmogorov-Smirnov test. For each pairwise cell type,
subclass, and gene the KS test was performed. Genes that were identified as 
significaltly (alpha = 0.05) different in all pairwise tests for a specific 
subclass were annotated as being specific for that class. A heatmap showing the
detected genes is shown below.

```{r kstest}
classHeatmap(geneList, counts.log, sampleClasses)
```
